---
title: "Assignment4_Group6"
author:
- familyname: Makornkhan
  othernames: Chatpisut
  address: Monash University
  email: cmak0010@student.monash.edu
  correspondingauthor: true
  qualifications:  Master of Business Analytics
- familyname: YANG
  othernames: BINGYU
  address: Monash University
  email: byan0033@student.monash.edu
  qualifications: Master of Business Analytics
- familyname: Trinh
  othernames: Phuong
  address: Monash University
  email: ptri0004@student.monash.edu
  correspondingauthor: true
  qualifications:  Master of Business Analytics
department: Faculty of \newline Business &\newline Economics
organization: Monash University
bibliography: references.bib
biblio-style: authoryear-comp
linestretch: 1.5
output:
  bookdown::pdf_document2:
    template: monashreport2.tex
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
    toc: true
---


```{r, echo = FALSE, message = FALSE, warning = FALSE}
# Libraries

library(tidyverse)
library(readr)
library(kableExtra)
library(bookdown)
library(dplyr)
```




# Report Introduction

The research project aims to explore and analyze the numbers and trends of human death caused by various air pollution.


# Section 1

```{r , echo = FALSE, message = FALSE, warning = FALSE}
Origin_data <- read_csv(here::here("Data/death-rates-from-air-pollution.csv"))
```

## Introduction

- @owidoutdoorairpollution point out that outdoor air pollution is one of the world's biggest health and environmental problems. The origin data in this section is from Our world in data and is based on the number of deaths caused by air pollution.

- This section is designed to research the four countries with the highest average number of deaths due to air pollution from 1990 to 2019 and to observe the changes in the number of deaths caused by the two main outdoor pollutants.


## Research question

Q1:Top four countries with the highest average number of deaths due to air pollution since 1990 to 2019.

Q2:Is air pollution in these four countries improving until 2019? 

Q3:The two main pollutants of outdoor air pollution are ozone and outdoor particulate matter. In the four countries with the highest average number of deaths caused by air pollution, what are the trends in the number of deaths caused by these two pollutants?

## Exploratory data analysis

- Q1

```{r , echo = FALSE}
Research_data <- Origin_data %>% 
  rename(Total_air_polution = 
           "Deaths - Cause: All causes - Risk: Air pollution - Sex: Both - Age: Age-standardized (Rate)",
         Indoor_air_pollution = 
           "Deaths - Cause: All causes - Risk: Household air pollution from solid fuels - Sex: Both - Age: Age-standardized (Rate)",
         Outdoor_particulate_matter = 
           "Deaths - Cause: All causes - Risk: Ambient particulate matter pollution - Sex: Both - Age: Age-standardized (Rate)",
         Outdoor_ozone_pollution = 
           "Deaths - Cause: All causes - Risk: Ambient ozone pollution - Sex: Both - Age: Age-standardized (Rate)")

```


```{r , echo = FALSE}
Q1_data <- Research_data %>% 
  group_by(Entity) %>% 
  summarise(mean_total_air_polution = mean(Total_air_polution,na.rm = TRUE)) %>% 
  arrange(desc(mean_total_air_polution)) %>% 
  head(4)

```


```{r airpollutiontab,echo=FALSE,message = FALSE, warning = FALSE}
kable(Q1_data, "latex", digits = 2,caption = "Average number of deaths from air pollution, 1990 to 2019", booktabs = T) %>% kable_styling(latex_options = c("striped", "hold_position", "scale_down"))
```

In table \@ref(tab:airpollutiontab), the four countries with the highest average number of deaths (per 100,000  population) due to air pollution are `r Q1_data$Entity[1]`,`r Q1_data$Entity[2]`,`r Q1_data$Entity[3]`,and `r Q1_data$Entity[4]`.

- Q2

```{r airpollutionfig, echo = FALSE,message = FALSE, warning = FALSE, fig.align = "center", fig.width = 6,fig.height=7, fig.cap = "Air pollution trends in different countries", fig.pos="H"}

Q2_data_1 <- Research_data %>% 
  filter(Entity %in% c("Solomon Islands","Afghanistan","Central African Republic","Somalia")) %>% 
  select(Entity,Year,Total_air_polution,Indoor_air_pollution)

Q2_figure_1 <- ggplot(Q2_data_1,aes(x = Year,
                     y = Total_air_polution,
                     color = Entity))+
  geom_line()+
  ylab("The number of deaths from air pollution")+
  ggtitle("Total air pollution")

Q2_figure_2 <- ggplot(Q2_data_1,aes(x = Year,
                     y = Indoor_air_pollution,
                     color = Entity))+
  geom_line()+
  ylab("The number of deaths from air pollution")+
  ggtitle("Indoor air pollution")

gridExtra::grid.arrange(Q2_figure_1,Q2_figure_2)
    
```


In figure \@ref(fig:airpollutionfig), I found significant improvements in total air pollution in these four countries, and a continuous downward trend in the number of deaths caused by indoor air pollution.

- Q3

```{r trendsfig, echo = FALSE,message = FALSE, warning = FALSE, fig.align = "center", fig.width = 6, fig.cap = "The trends in the number of deaths caused by these two pollutants", fig.pos="H"}
Q2_data_2 <- Research_data %>% 
 filter(Entity %in% c("Solomon Islands","Afghanistan","Central African Republic","Somalia")) %>% 
  select(Entity,Year,Outdoor_particulate_matter, Outdoor_ozone_pollution) %>% 
   pivot_longer(cols = c(Outdoor_particulate_matter, Outdoor_ozone_pollution),
              names_to = "Air_pollution_type",
              values_to = "The_number_of_deaths")

ggplot(Q2_data_2,aes(x = Year,
                   y = The_number_of_deaths,
                   color = Air_pollution_type))+
  geom_line()+
  facet_wrap(~Entity,nrow = 4,scales = "free")+
  ylab("The number of deaths (per 100,000 ppl)")
  
```

In figure \@ref(fig:trendsfig), I found a small increase in the number of deaths caused by particulate pollution in Solomon Islands,Central African Republic and Afghanistan, while only a very small number of people in Somalia die from outdoor pollution.


## Conclusion 

From the above studies, I conclude that the number of deaths due to indoor air pollution has declined in most countries in recent decades. However, the number of deaths due to outdoor air pollution, such as outdoor particulate matter, is still on the rise. Therefore, @neidell2004air indicated that the world should regularly monitor air quality and reduce sources of particulate pollution.

\section*{Vietnam and New Zealand}

# Section 2

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.pos = 'H')
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
death <- read_csv(here::here("Data/death-rates-from-air-pollution.csv"))
```

## Introduction

Pollution is not only the leading factor taking away peoples’ lives, it also leaves serious long-term effects on our living quality. 

In this section we will compare the relationship between death rates and different types of pollution from 1990 to 2019 in two countries, Vietnam and New Zealand.

*The data set was originated from @owidairpollution.*

```{r,  echo = FALSE, message = FALSE, warning = FALSE}
death_rate <- death %>%
  filter(Entity %in% c("Vietnam", "New Zealand")) %>%
  rename(Household_air_pollution = "Deaths - Cause: All causes - Risk: Household air pollution from solid fuels - Sex: Both - Age: Age-standardized (Rate)",
         Ambient_particulate_matter_pollution = "Deaths - Cause: All causes - Risk: Ambient particulate matter pollution - Sex: Both - Age: Age-standardized (Rate)",
         Air_pollution = "Deaths - Cause: All causes - Risk: Air pollution - Sex: Both - Age: Age-standardized (Rate)",
         Ambient_ozone_pollution = "Deaths - Cause: All causes - Risk: Ambient ozone pollution - Sex: Both - Age: Age-standardized (Rate)") %>%
  pivot_longer(cols = c("Household_air_pollution",
               "Ambient_particulate_matter_pollution",
               "Air_pollution", 
               "Ambient_ozone_pollution"),
               names_to = "Risk_factor",
               values_to = "Rate")
```

## Research question

- Q1: Which type of pollution is the most common attribute of death in each country?

- Q2: How did the death rates associated with different types of pollution change over the years in both countries?

## Exploratory data analysis

- Q1

From table \@ref(tab:nz), we can observe that air pollution contributed most to the death rates in New Zealand with 7 deaths per 100,000 people.

```{r,  echo = FALSE, message = FALSE, warning = FALSE}
death_cause <- death_rate %>%
  group_by(Entity, Risk_factor) %>%
  summarise(mean_rate = mean(Rate))
  
death_cause$mean_rate <- round(death_cause$mean_rate, digits = 2)
  
```

```{r nz,  echo = FALSE, message=FALSE, warning=FALSE, fig.pos='H'}
nz <- death_cause %>%
  filter(Entity == "New Zealand")

kable(nz, "latex", caption = "Death rate related to different risk factors in New Zealand", booktabs = T) %>%
kable_styling(latex_options = c('striped', 'hold_position'))
```




```{r vn,  echo = FALSE, message = FALSE, warning = FALSE, fig.pos='H'}
vn <- death_cause %>%
  filter(Entity == "Vietnam")

kable(vn, "latex", caption = "Death rate related to different risk factors in Vietnam", booktabs = T) %>% kable_styling(latex_options = c('striped', 'hold_position'))
```

Similarly, in table \@ref(tab:vn), it can be seen that air pollution is also the leading factor in Vietnam, accounting for 132 deaths per 100,000 people.


According to @mannucci2017health, there were almost seven million deaths associated with the effects of air pollution. In addition, the impacts tend to be greater across low and middle income countries, mainly due to the increasing shift to industrialization. The notion has been demonstrated in the case of Vietnam and New Zealand, where the number of deaths in Vietnam is approximately 18 times higher than New Zealand.

- Q2

The graph \@ref(fig:nzfig) shows the rate at which people died due to the effects of different types of pollution over a 29-year period from 1990 to 2019. 

There was an overall downward trend in the impacts of air pollution and ambient particulate matter pollution to death rate. 
Rate of people died associated air pollution originated from household and ambient ozone pollution showed a slight decrease over the years.

In contrast, the rate at which pollution from ambient ozone contributed to death remained relatively steady.

```{r nzfig,  echo = FALSE, message = FALSE, warning = FALSE, out.width='70%', fig.cap="Death rates attributed to pollution in New Zealand from 1990 to 2019", fig.pos= 'H', fig.align='center'}
nz_rate <- death_rate %>%
  filter(Entity == "New Zealand")

ggplot(nz_rate, aes(x = Year, y = Rate, color = Risk_factor)) +
  geom_line() +
  scale_x_continuous(breaks = seq(1990, 2019, by=2)) +
  labs(y = "Death rate",
       color = "Risk factor") +
  theme(axis.text.x = element_text(angle=45, hjust = 1))

```
Overall, in \@ref(fig:vnfig), death rate associated with air pollution and pollution resulted from household decreased dramatically over the years. 

However, the rate attributable to pollution from ambient particulate matter pollution showed a noticeable surge despite having a slight drop in 2010.

Similarly, deaths impacted by pollution from ambient ozone contributed to global death showed little or no signs of changing during the period.


```{r vnfig,  echo = FALSE, message = FALSE, warning = FALSE, out.width='70%', fig.cap="Death rates attributed to pollution in Vietnam from 1990 to 2019", fig.pos= 'H', fig.align='center'}
vn_rate <- death_rate %>%
  filter(Entity == "Vietnam")

ggplot(vn_rate, aes(x = Year, y = Rate, color = Risk_factor)) +
  geom_line() +
  scale_x_continuous(breaks = seq(1990, 2019, by=2)) +
  labs(y = "Death rate",
       color = "Risk factor") +
  labs(y = "Death rate") +
  theme(axis.text.x = element_text(angle=45, hjust = 1))
```



## Conclusion

Overall, it is obvious that air pollution is the largest contributor of deaths in both countries.

It can be seen that there was a overall decrease in deaths attributed to air pollution and pollution from household in recent decades. 

However, the death rate in Vietnam remained much higher than in New Zealand. @gordon2014respiratory explained this by the increase in urbanization and lack of access to clean fuels for cooking in developing nations.


# Section 3

## Introduction


Last but not least, we would like to further explore countries with extreme values on both ends. Henceforth, we can see the nature of trends and numbers across the year, along with reasonable research and facts which support those unfortunate number of deaths by pollution. Which lead us to formulate research questions as below:


## Research Questions

- Q1: Which country has the highest number of Air Pollution caused casualties? And what is its trend across the year pertaining to their cause of deaths severity?
- Q2: What is the 4 lowest polluted countries? Regarding the comparison between them, what are possible reasons for these significantly low results?

```{r read data, echo = FALSE, message = FALSE, warning = FALSE}
data <- read.csv("Data/death-rates-from-air-pollution.csv")
```

```{r clean data 1, echo = FALSE, message = FALSE, warning = FALSE}
air_pollution <- data %>% 
  rename(Country = Entity) %>% 
  rename(Deaths_From_Solid_Fuels = 
         Deaths...Cause..All.causes...Risk..Household.air.pollution.from.solid.fuels...Sex..Both...Age..Age.standardized..Rate.) %>%
  rename(Deaths_From_Ambient_Matter =
         Deaths...Cause..All.causes...Risk..Ambient.particulate.matter.pollution...Sex..Both...Age..Age.standardized..Rate.) %>% 
  rename(Deaths_From_Air_Pollution = Deaths...Cause..All.causes...Risk..Air.pollution...Sex..Both...Age..Age.standardized..Rate.) %>% 
  rename(Deaths_Ambient_Ozone = Deaths...Cause..All.causes...Risk..Ambient.ozone.pollution...Sex..Both...Age..Age.standardized..Rate.) %>% 
  select("Country", "Year", "Deaths_From_Solid_Fuels", "Deaths_From_Ambient_Matter", "Deaths_From_Air_Pollution", "Deaths_Ambient_Ozone")
```


## Exploratory data analysis

By exploring the highest ends of the data, we discovered that Solomon Islands has the greatest number of deaths caused by pollution in significant margins of gap comparatively to latter ranked countries. Therefore, Solomon Islands data set is further explored into detail on its trend across the year.

```{r Question 1, echo = FALSE, message = FALSE, warning = FALSE}
air_pollution_highest <- air_pollution %>% 
  select("Country", "Year", "Deaths_From_Air_Pollution") %>% 
  group_by(Country) %>%
  summarise(Deaths_From_Air_Pollution = mean(Deaths_From_Air_Pollution, na.rm=TRUE)) %>%
  arrange(desc(Deaths_From_Air_Pollution))

air_pollution_SI <- air_pollution %>% 
  filter(Country == "Solomon Islands")
```


```{r figure1, echo = FALSE, message = FALSE, warning = FALSE, fig.align='center', fig.cap = "Solomon Islands - Annual Trend", out.width='80%'}
colors <- c("Deaths From Air Pollution" = "red", "Deaths From Solid Fuels" = "blue", "Deaths From Ambient Matter" = "green", "Deaths Ambient Ozone" = "yellow")

air_pollution_SI %>%
  ggplot(aes(x = Year)) +
  geom_line(aes(y = Deaths_From_Air_Pollution, color = "Deaths From Air Pollution")) +
  geom_line(aes(y = Deaths_From_Solid_Fuels, color = "Deaths From Solid Fuels")) +
  geom_line(aes(y = Deaths_From_Ambient_Matter, color = "Deaths From Ambient Matter")) +
  geom_line(aes(y = Deaths_Ambient_Ozone, color = "Deaths Ambient Ozone")) +
  labs(y = "Death Rate per Type of Pollution", color = "Legend") +
  scale_color_manual(values = colors)
```

Referring to figure \@ref(fig:figure1)) above, we can see that, although trends are in the declining movement since the start of 1990, the number of deaths caused by both air pollution and solid fuels are terrifying high comparatively. With these numbers, Solomon Islands is assured to be in the stage of air pollution crisis. 

As stated by @hunt2011policy in the research paper, health hazards endpoints which quantified are including, but not limited to Premature Mortality, Respiratory and Cardiovascular Illness, Cancer, or even Infant Mortality. Therefore, Solomon Islands living condition is still at an alarming stage which need immediate addressed.


```{r Question 2, echo = FALSE, message = FALSE, warning = FALSE, fig.pos="H"}
air_pollution_lowest <- air_pollution %>% 
  select("Country", "Year", "Deaths_From_Air_Pollution") %>% 
  group_by(Country) %>%
  summarise(Deaths_From_Air_Pollution = mean(Deaths_From_Air_Pollution, na.rm=TRUE)) %>%
  arrange(Deaths_From_Air_Pollution)

air_pollution_lowest_cnty <- air_pollution %>% 
  select("Country", "Year", "Deaths_From_Air_Pollution") %>% 
  group_by(Country) %>% 
  filter(Country == c("Iceland", "New Zealand", "Australia", "Puerto Rico")) %>% 
  summarise(Deaths_From_Air_Pollution = mean(Deaths_From_Air_Pollution, na.rm=TRUE)) %>%
  arrange(Deaths_From_Air_Pollution) 

air_pollution_lowest_other <- air_pollution %>% 
  select("Country", "Year", "Deaths_From_Solid_Fuels") %>% 
  group_by(Country) %>% 
  filter(Country == c("Iceland", "New Zealand", "Australia", "Puerto Rico")) %>% 
  summarise(Deaths_From_Solid_Fuels = mean(Deaths_From_Solid_Fuels, na.rm=TRUE)) %>%
  arrange(Deaths_From_Solid_Fuels) 
```

Following this, we are looking at the lowest end for least number of deaths caused by pollution; in which Iceland, New Zealand, Australia, and Puerto Rico hold top spots.  


```{r table1, echo = FALSE, message = FALSE, warning = FALSE, fig.align='center', fig.cap = "Lowest Air Pollution based on Countries", fig.pos="H"}
air_pollution_lowest_cnty %>%
  kable(caption = "Top 4 Countries with lowest Air Pollution Death", digits=3) %>% 
  kable_styling(latex_options = "HOLD_position")
```

From table \@ref(tab:table1)), we can see that less than 9 persons on average resulted in a fatal incident caused by air pollution.

```{r table2, echo = FALSE, message = FALSE, warning = FALSE, fig.align='center', fig.cap = "Lowest Air Pollution based on Countries", fig.pos="H"}
air_pollution_lowest_other %>%
  kable(caption = "Top 4 Countries with lowest Air Pollution Death from Solid Fuels", digits=3) %>% 
  kable_styling(latex_options = "HOLD_position")
```

Furthermore, comparatively to solid fuels-led death, with even less numbers are shown as a result in table \@ref(tab:table2)).

By using Iceland as the main example, according to @barsotti2020probabilistic, even though air pollution rate is generally excellent in Iceland, a subtle surge in chemical or pollution such as SO2 is considered as a threat to their ozone and breathing environment. Which they have immediate action to tackle and resolve problem as quick as possible. 


## Conclusion

Hence, these countries are ensured that rapid response in air pollution issue could make a potential bright further for their countries' air for generations.

# Report Conclusion

In conclusion, we can see that air pollution is subtle yet highly severe global issues. Even though many countries can manage to suppress their toxic footprint, this is a global matter that every country needs to address for the better and brighter future of air pollution-less world.







