---
title: "Assignment4_Group6"
author:
- familyname: Makornkhan
  othernames: Chatpisut
  address: Monash University
  email: cmak0010@student.monash.edu
  correspondingauthor: true
  qualifications:  Master of Business Analytics
- familyname: YANG
  othernames: BINGYU
  address: Monash University
  email: byan0033@student.monash.edu
  qualifications: Master of Business Analytics
- familyname: Trinh
  othernames: Phuong
  address: Monash University
  email: ptri0004@student.monash.edu
  correspondingauthor: true
  qualifications:  Master of Business Analytics
department: Faculty of \newline Business &\newline Economics
organization: Monash University
bibliography: references.bib
biblio-style: authoryear-comp
linestretch: 1.5
output:
  bookdown::pdf_document2:
    template: monashreport2.tex
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
    toc: true
---


```{r, echo = FALSE, message = FALSE, warning = FALSE}
# Libraries

library(tidyverse)
library(readr)
library(kableExtra)
library(bookdown)
```
\section*{Vietnam and New Zealand}

# Section 2


```{r}
death <- read_csv(here::here("Data/death-rates-from-air-pollution.csv"))
```
# Introduction:

Pollution is not only the leading factor that takes away peoples’ lives, it also also leaves considerable effects on our living quality. 

In this section we will compare the relationship between death rates and different types of pollution from 1990 to 2019 in two countries, Vietnam and New Zealand.

*The data set was originated from @owidairpollution.*

```{r}
death_rate <- death %>%
  filter(Entity %in% c("Vietnam", "New Zealand")) %>%
  rename(Household_air_pollution = "Deaths - Cause: All causes - Risk: Household air pollution from solid fuels - Sex: Both - Age: Age-standardized (Rate)",
         Ambient_particulate_matter_pollution = "Deaths - Cause: All causes - Risk: Ambient particulate matter pollution - Sex: Both - Age: Age-standardized (Rate)",
         Air_pollution = "Deaths - Cause: All causes - Risk: Air pollution - Sex: Both - Age: Age-standardized (Rate)",
         Ambient_ozone_pollution = "Deaths - Cause: All causes - Risk: Ambient ozone pollution - Sex: Both - Age: Age-standardized (Rate)") %>%
  pivot_longer(cols = c("Household_air_pollution",
               "Ambient_particulate_matter_pollution",
               "Air_pollution", 
               "Ambient_ozone_pollution"),
               names_to = "Risk_factor",
               values_to = "Rate")
```
















