---
title: "Assignment4_Group6"
author:
- familyname: Makornkhan
  othernames: Chatpisut
  address: Monash University
  email: cmak0010@student.monash.edu
  correspondingauthor: true
  qualifications:  Master of Business Analytics
- familyname: YANG
  othernames: BINGYU
  address: Monash University
  email: byan0033@student.monash.edu
  qualifications: Master of Business Analytics
- familyname: Trinh
  othernames: Phuong
  address: Monash University
  email: ptri0004@student.monash.edu
  correspondingauthor: true
  qualifications:  Master of Business Analytics
department: Faculty of \newline Business &\newline Economics
organization: Monash University
bibliography: references.bib
biblio-style: authoryear-comp
linestretch: 1.5
output:
  bookdown::pdf_document2:
    template: monashreport2.tex
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
    toc: true
---


```{r, echo = FALSE, message = FALSE, warning = FALSE}
# Libraries

library(tidyverse)
library(readr)
library(kableExtra)
library(bookdown)
library(dplyr)

```

## Section 3

### Introduction

\section*{Section 3 - Highest and Lowest Air Pollution}


**Research Questions**

Last but not least, we would like to further explore countries with extreme values on both ends. Henceforth, we can see the nature of trends and numbers across the year, along with reasonable research and facts which support those unfortunate number of deaths by pollution. Which lead us to formulate research questions as below:

- Q1: Which country has the highest number of Air Pollution caused casualties? And what is its trend across the year pertaining to their cause of deaths severity?
- Q2: What is the 4 lowest polluted countries? Regarding the comparison between them, what are possible reasons for these significantly low results?

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.pos = 'H')
```

```{r read data, echo = FALSE, message = FALSE, warning = FALSE}
data <- read.csv("Data/death-rates-from-air-pollution.csv")
```

```{r clean data 1, echo = FALSE, message = FALSE, warning = FALSE}
air_pollution <- data %>% 
  rename(Country = Entity) %>% 
  rename(Deaths_From_Solid_Fuels = 
         Deaths...Cause..All.causes...Risk..Household.air.pollution.from.solid.fuels...Sex..Both...Age..Age.standardized..Rate.) %>%
  rename(Deaths_From_Ambient_Matter =
         Deaths...Cause..All.causes...Risk..Ambient.particulate.matter.pollution...Sex..Both...Age..Age.standardized..Rate.) %>% 
  rename(Deaths_From_Air_Pollution = Deaths...Cause..All.causes...Risk..Air.pollution...Sex..Both...Age..Age.standardized..Rate.) %>% 
  rename(Deaths_Ambient_Ozone = Deaths...Cause..All.causes...Risk..Ambient.ozone.pollution...Sex..Both...Age..Age.standardized..Rate.) %>% 
  select("Country", "Year", "Deaths_From_Solid_Fuels", "Deaths_From_Ambient_Matter", "Deaths_From_Air_Pollution", "Deaths_Ambient_Ozone")
```


### Analysis

By exploring the highest ends of the data, we discovered that Solomon Islands has the greatest number of deaths caused by pollution in significant margins of gap comparatively to latter ranked countries. Therefore, Solomon Islands data set is further explored into detail on its trend across the year.

```{r Question 1, echo = FALSE, message = FALSE, warning = FALSE}
air_pollution_highest <- air_pollution %>% 
  select("Country", "Year", "Deaths_From_Air_Pollution") %>% 
  group_by(Country) %>%
  summarise(Deaths_From_Air_Pollution = mean(Deaths_From_Air_Pollution, na.rm=TRUE)) %>%
  arrange(desc(Deaths_From_Air_Pollution))

air_pollution_SI <- air_pollution %>% 
  filter(Country == "Solomon Islands")
```


```{r figure1, echo = FALSE, message = FALSE, warning = FALSE, fig.align='center', fig.cap = "Solomon Islands - Annual Trend", out.width='80%'}
colors <- c("Deaths From Air Pollution" = "red", "Deaths From Solid Fuels" = "blue", "Deaths From Ambient Matter" = "green", "Deaths Ambient Ozone" = "yellow")

air_pollution_SI %>%
  ggplot(aes(x = Year)) +
  geom_line(aes(y = Deaths_From_Air_Pollution, color = "Deaths From Air Pollution")) +
  geom_line(aes(y = Deaths_From_Solid_Fuels, color = "Deaths From Solid Fuels")) +
  geom_line(aes(y = Deaths_From_Ambient_Matter, color = "Deaths From Ambient Matter")) +
  geom_line(aes(y = Deaths_Ambient_Ozone, color = "Deaths Ambient Ozone")) +
  labs(y = "Death Rate per Type of Pollution", color = "Legend") +
  scale_color_manual(values = colors)
```

Referring to figure\@ref(fig:figure1)) above, we can see that, although trends are in the declining movement since the start of 1990, the number of deaths caused by both air pollution and solid fuels are terrifying high comparatively. With these numbers, Solomon Islands is assured to be in the stage of air pollution crisis. 

As stated by @hunt2011policy in the research paper, health hazards endpoints which quantified are including, but not limited to Premature Mortality, Respiratory and Cardiovascular Illness, Cancer, or even Infant Mortality. Therefore, Solomon Islands living condition is still at an alarming stage which need immediate addressed.


```{r Question 2, echo = FALSE, message = FALSE, warning = FALSE, fig.pos="H"}
air_pollution_lowest <- air_pollution %>% 
  select("Country", "Year", "Deaths_From_Air_Pollution") %>% 
  group_by(Country) %>%
  summarise(Deaths_From_Air_Pollution = mean(Deaths_From_Air_Pollution, na.rm=TRUE)) %>%
  arrange(Deaths_From_Air_Pollution)

air_pollution_lowest_cnty <- air_pollution %>% 
  select("Country", "Year", "Deaths_From_Air_Pollution") %>% 
  group_by(Country) %>% 
  filter(Country == c("Iceland", "New Zealand", "Australia", "Puerto Rico")) %>% 
  summarise(Deaths_From_Air_Pollution = mean(Deaths_From_Air_Pollution, na.rm=TRUE)) %>%
  arrange(Deaths_From_Air_Pollution) 

air_pollution_lowest_other <- air_pollution %>% 
  select("Country", "Year", "Deaths_From_Solid_Fuels") %>% 
  group_by(Country) %>% 
  filter(Country == c("Iceland", "New Zealand", "Australia", "Puerto Rico")) %>% 
  summarise(Deaths_From_Solid_Fuels = mean(Deaths_From_Solid_Fuels, na.rm=TRUE)) %>%
  arrange(Deaths_From_Solid_Fuels) 
```

Following this, we are looking at the lowest end for least number of deaths caused by pollution; in which Iceland, New Zealand, Australia, and Puerto Rico hold top spots.  


```{r table1, echo = FALSE, message = FALSE, warning = FALSE, fig.align='center', fig.cap = "Lowest Air Pollution based on Countries", fig.pos="H"}
air_pollution_lowest_cnty %>%
  kable(caption = "Top 4 Countries with lowest Air Pollution Death", digits=3) %>% 
  kable_styling(latex_options = "HOLD_position")
```

From table\@ref(tab:table1)), we can see that less than 9 persons on average resulted in a fatal incident caused by air pollution.

```{r table2, echo = FALSE, message = FALSE, warning = FALSE, fig.align='center', fig.cap = "Lowest Air Pollution based on Countries", fig.pos="H"}
air_pollution_lowest_other %>%
  kable(caption = "Top 4 Countries with lowest Air Pollution Death from Solid Fuels", digits=3) %>% 
  kable_styling(latex_options = "HOLD_position")
```

Furthermore, comparatively to solid fuels-led death, with even less numbers are shown as a result in table \@ref(tab:table2)).

By using Iceland as the main example, according to @barsotti2020probabilistic, even though air pollution rate is generally excellent in Iceland, a subtle surge in chemical or pollution such as SO2 is considered as a threat to their ozone and breathing environment. Which they have immediate action to tackle and resolve problem as quick as possible. 

Hence, these countries are ensured that rapid response in air pollution issue could make a potential bright further for their countries' air for generations.


In conclusion, 


